The Midplane is Not a Plane: Implications for {\em Gaia}

### 2019-03-25 paper/ms.tex
Fix indentation throughout

\section{Introduction} \label{sec:intro}
Our understanding of the Milky Way is currently undergoing a revolution in the
era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
major merger
\citep{2018ApJ...860L..11K,2018Natur.563...85H,2018arXiv180704290L,2019MNRAS.482.3426M},
a phase-space ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly indicating either local
substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L}, bar
buckling \citep{2018arXiv181109205K}, although may just be contributions from moving groups \citep{2019arXiv190308325M} and a gap suggestive of perturbation by a dark matter
subtructure in the tidal stream GD1 \citep{2018arXiv181103631B}.
These discoveries all indicate that the Milky Way's stellar distribution, which demonstrably departs significantly from axisymmetry, is undergoing phase mixing and dynamical interactions across a range of spatial and temporal scales.

@@

\begin{figure*}
\plotone{fig/midplane_fit.pdf}
\caption{The local midplane determined at the fiducial Solar radius
($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} ({\em left},
{\em center}, and {\em right} panels). The local midplane is determined at a
position $\phi$ by taking the median height of all stars within $R=0.5\,\kpc$
and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
using the new height $10$ times to converge on the local midplane height. In
order to allow for the possibility that the fiducial Galactocentric coordinate
system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
each panel --- this figure is reproduced with the original midplane
determination (i.e. before subtracting the best fit sine curve) in
Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars
within a $2\,\kpc$ height of the fiducial midplane to determine error bars
(95\uth and 5\uth percentiles), which we report as dashed lines.}
\label{fig:midplane}
\end{figure*}

@@

%%%move this later - decide where, probably conclusions
%While we focus our discussion now solely on the midplane, there is an
%additional complication in the measurement of the distance from the Sun to the
%Galactic center that may also be important. When modelling stars at Galactic
%radii close to the Sun's radius, it is in principle possible to determine what
%point they are orbiting about: the ``dynamical Galactic
%center.'' It is not guaranteed that this dynamical center will coincide
%with the distance to Sgr~A\textsuperscript{*} \citep{2018AA...615L..15G}. In
%fact, \citet{2015ApJ...803...80K} were able to measure the distance of the Sun
%to the Galactic center using the dynamical information of Palomar 5. Other
%dynamical measurements of the Solar distance have been made, though none with
%a precision capable of competing with the distance to Sgr~A\textsuperscript{*}
%\citep{1981gask.book.....M,2011PASJ...63..867S,2012MNRAS.427..274S,2013AstL...39...95B,2013IAUS..289..444Z}.

>>>>

\section{Introduction} \label{sec:intro}
Our understanding of the Milky Way is currently undergoing a revolution in the
era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
major merger \citep{2018ApJ...860L..11K, 2018Natur.563...85H,
2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the
solar neighborhood \citep{2018Natur.561..360A} possibly indicating either
local substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L}, bar
buckling \citep{2018arXiv181109205K}, although may just be contributions from
moving groups \citep{2019arXiv190308325M} and a gap suggestive of perturbation
by a dark matter subtructure in the tidal stream GD1
\citep{2018arXiv181103631B}. These discoveries all indicate that the Milky
Way's stellar distribution, which demonstrably departs significantly from
axisymmetry, is undergoing phase mixing and dynamical interactions across a
range of spatial and temporal scales.

@@

\begin{figure*}
\plotone{fig/midplane_fit.pdf}
\caption{The local midplane determined at the fiducial Solar radius
($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} ({\em left},
{\em center}, and {\em right} panels). The local midplane is determined at a
position $\phi$ by taking the median height of all stars within $R=0.5\,\kpc$
and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
using the new height $10$ times to converge on the local midplane height. In
order to allow for the possibility that the fiducial galactocentric coordinate
system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
each panel --- this figure is reproduced with the original midplane
determination (i.e. before subtracting the best fit sine curve) in
Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars
within a $2\,\kpc$ height of the fiducial midplane to determine error bars
(95\uth and 5\uth percentiles), which we report as dashed lines.}
\label{fig:midplane}
\end{figure*}

@@

%%%move this later - decide where, probably conclusions While we focus our
%discussion now solely on the midplane, there is an additional complication in
%the measurement of the distance from the Sun to the Galactic center that may
%also be important. When modelling stars at Galactic radii close to the Sun's
%radius, it is in principle possible to determine what point they are orbiting
%about: the ``dynamical Galactic center.'' It is not guaranteed that this
%dynamical center will coincide with the distance to Sgr~A\textsuperscript{*}
%\citep{2018AA...615L..15G}. In fact, \citet{2015ApJ...803...80K} were able to
%measure the distance of the Sun to the Galactic center using the dynamical
%information of Palomar 5. Other dynamical measurements of the Solar distance
%have been made, though none with a precision capable of competing with the
%distance to Sgr~A\textsuperscript{*}
%\citep{1981gask.book.....M,2011PASJ...63..867S,2012MNRAS.427..274S,2013AstL...39...95B,2013IAUS..289..444Z}.
